---
title: "An example of spatial incidence data"
author: "Thibaut Jombart"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---



```{r, echo = FALSE}
options(width = 100)

knitr::opts_chunk$set(
  fig.width = 7, 
  fig.height = 7,
  dev = 'png',
  fig.path = "figs/"
)
```




```{r load_pkg}

library(outbreaks)
library(ggmap)
library(ggplot2)

```


## Toy example: simulated Ebola outbreak in Freetown

This example uses the latest devel version of [http://www.repidemicsconsortium.org/outbreaks/](*outbreaks*). 

```{r data}

head(ebola_sim$linelist)
lonlat <- ebola_sim$linelist[, 10:11]
head(lonlat)
nrow(lonlat)

```

We assume latitude / longitude will be the standard way to indicate locations. A
basic map, using *ggmap* and *ggplot2*, would look like:

```{r basic_map}

base_points <- ggplot(data = lonlat, aes(x = lon, y = lat))
base <- ggmap(get_map(lonlat, zoom = 13), base_layer = base_points)
base + geom_point(col = "red", alpha = .2)

```

Adding kernel density estimation would also make sense here:
```{r density_map}

base + geom_point(col = "red", alpha = .2) + geom_density_2d()

```

It would also make sense to aggregate data first. This is a bit less trivial as
there are several options for this:

- use a clustering method based on lon/lat, and defining locations as geographic
  centres of the clusters

- round up locations and use `xyTable`

We illustrate the latter, which may be more likely to reflect actual situations
where locations have low resolution (e.g. village, district level, etc.).

```{r aggregate_map}

xy_rough <- data.frame(xyTable(lonlat, digits = 4))
base + geom_point(data = xy_rough, aes(x = x, y = y, size = number),
       		  col = "red", alpha = .5)

```

And here is the clustering version; we use $k-means$ as it is fairly quick and
offers the nice feature of the underlying Voronoi tesselation space
partitioning:

```{r clust_map}

clust <- kmeans(lonlat, centers = 60)
xy_clust <- data.frame(clust$centers, number = clust$size)
base + geom_point(data = xy_clust, aes(x = lon, y = lat, size = number),
       		  col = "red", alpha = .5)

```